---
title: "Marketing Project: Dal Rae Project"
author: "Landon Smith"
date: "`r Sys.Date()`"
output: word_document
---
```{r}
library(dplyr)

#Import Sales Data
sales_2018 = read.csv("/Users/landonsmith/Desktop/Dal Rae CSVs/2018 Sales.csv")
sales_2019 = read.csv("/Users/landonsmith/Desktop/Dal Rae CSVs/2019 Sales.csv")
sales_2020 = read.csv("/Users/landonsmith/Desktop/Dal Rae CSVs/2020 Sales.csv")
sales_2021 = read.csv("/Users/landonsmith/Desktop/Dal Rae CSVs/2021 Sales.csv")
sales_2022 = read.csv("/Users/landonsmith/Desktop/Dal Rae CSVs/2022 Sales.csv")
sales_2023 = read.csv("/Users/landonsmith/Desktop/Dal Rae CSVs/2023 Sales.csv")

#Drop Unnecessary Columns
sales_2018 = subset(sales_2018, select = c(Item.Num, Item.Name, Num.Sold, Price.Sold, Amount))
sales_2019 = subset(sales_2019, select = c(Item.Num, Item.Name, Num.Sold, Price.Sold, Amount))
sales_2020 = subset(sales_2020, select = c(Item.Num, Item.Name, Num.Sold, Price.Sold, Amount))
sales_2021 = subset(sales_2021, select = c(Item.Num, Item.Name, Num.Sold, Price.Sold, Amount))
sales_2022 = subset(sales_2022, select = c(Item.Num, Item.Name, Num.Sold, Price.Sold, Amount))
sales_2023 = subset(sales_2023, select = c(Item.Num, Item.Name, Num.Sold, Price.Sold, Amount))

#Change Columns From Character Datatype to Numeric
sales_2018$Num.Sold = as.numeric(sales_2018$Num.Sold)
sales_2018$Price.Sold = as.numeric(sales_2018$Price.Sold)

sales_2019$Num.Sold = as.numeric(sales_2019$Num.Sold)
sales_2019$Price.Sold = as.numeric(sales_2019$Price.Sold)

sales_2020$Num.Sold = as.numeric(sales_2020$Num.Sold)
sales_2020$Price.Sold = as.numeric(sales_2020$Price.Sold)

sales_2021$Num.Sold = as.numeric(sales_2021$Num.Sold)
sales_2021$Price.Sold = as.numeric(sales_2021$Price.Sold)

sales_2022$Num.Sold = as.numeric(sales_2022$Num.Sold)
sales_2022$Price.Sold = as.numeric(sales_2022$Price.Sold)

sales_2023$Num.Sold = as.numeric(sales_2023$Num.Sold)
sales_2023$Price.Sold = as.numeric(sales_2023$Price.Sold)

#Combine all Datasets
sales_all_years <- rbind(sales_2018, sales_2019, sales_2020, sales_2021, sales_2022, sales_2023)
#sales_all_years <- sales_2023
sales_all_years

```

### Calculate Total Number Sold at Each Pricepoint
```{r}
#R is case sensitive, query by upper and lower case

#Low Tier: Poor Man's Pepper Steak 
poor_mans <- subset(sales_all_years, grepl("Poor", Item.Name))
poor_mans

poor_mans_grouped = poor_mans %>% group_by(Price.Sold) %>% summarise(number_sold = sum(Num.Sold))
poor_mans_grouped

#Medium Tier: Medallions of Filet Mignon
medallions <- subset(sales_all_years, grepl("Med", Item.Name))
medallions

medallions_grouped = medallions %>% group_by(Price.Sold) %>% summarise(number_sold = sum(Num.Sold))
medallions_grouped

#High Tier: Dal Rae Famous Pepper Steak 12oz
famous_pepper <- subset(sales_all_years, grepl("Peppr", Item.Name))
famous_pepper

famous_pepper_grouped = famous_pepper %>% group_by(Price.Sold) %>% summarise(number_sold = sum(Num.Sold))
famous_pepper_grouped
```

### Removes pricepoints with volume too low to qualify & bundles
```{r}
#Low Tier: Poor Man's Pepper Steak 
poor_mans_filtered = poor_mans_grouped[poor_mans_grouped$number_sold > 100, ]
poor_mans_filtered

#Medium Tier: Medallions of Filet Mignon
medallions_filtered = medallions_grouped[medallions_grouped$number_sold > 300, ]
medallions_filtered

med_number_sold_vector = medallions_filtered$number_sold
med_price_sold_vector = medallions_filtered$Price.Sold
med_number_sold_vector
med_price_sold_vector

#High Tier: Dal Rae Famous Pepper Steak 12oz
famous_pepper_filtered = famous_pepper_grouped[famous_pepper_grouped$number_sold > 300, ]

#Removes Bundle Prices from Low/High Prices
poor_mans_filtered = poor_mans_filtered[poor_mans_filtered$Price.Sold > 9, ]
poor_mans_filtered

low_number_sold_vector = poor_mans_filtered$number_sold
low_price_sold_vector = poor_mans_filtered$Price.Sold
low_number_sold_vector
low_price_sold_vector

famous_pepper_filtered = famous_pepper_filtered[famous_pepper_filtered$Price.Sold < 99, ]
famous_pepper_filtered

high_number_sold_vector = famous_pepper_filtered$number_sold
high_price_sold_vector = famous_pepper_filtered$Price.Sold
high_number_sold_vector
high_price_sold_vector
```
### Import Manually Collected Data - Days on the Menu
```{r}
library(readxl)

days_on_menu_poor_mans <- read_excel("~/Desktop/Dal Rae CSVs/Price Tier Days on Menu.xlsx", sheet = "Poor Man's")
days_on_menu_medallions <- read_excel("~/Desktop/Dal Rae CSVs/Price Tier Days on Menu.xlsx", sheet = "Medallions")
days_on_menu_famous_pepper <- read_excel("~/Desktop/Dal Rae CSVs/Price Tier Days on Menu.xlsx", sheet = "12oz Pepper")

poor_mans_filtered$days_on_menu = days_on_menu_poor_mans$days_on_menu
poor_mans_filtered

medallions_filtered$days_on_menu = days_on_menu_medallions$days_on_menu
medallions_filtered

famous_pepper_filtered$days_on_menu = days_on_menu_famous_pepper$days_on_menu
famous_pepper_filtered
```
### Intermediate Calculation for SLide
```{r}
poor_man_average = mean(poor_mans_filtered$Price.Sold)
medallions_average = mean(medallions_filtered$Price.Sold)
pepper_average = mean(famous_pepper_filtered$Price.Sold)

poor_man_sum = sum(poor_mans_filtered$number_sold)
medallions_sum = sum(medallions_filtered$number_sold)
pepper_sum = sum(famous_pepper_filtered$number_sold)

entree = c("Poor Man's Pepper Steak ($20-$30)", "Medallions of Filet Mignon ($30-$50)", "Dal Rae Famous Pepper Steak ($50+)")
avg_price_sold = c(poor_man_average, medallions_average, pepper_average)
number_sold = c(poor_man_sum, medallions_sum, pepper_sum)

summary = data.frame(entree, avg_price_sold, number_sold)
summary$revenue = summary$avg_price_sold * summary$number_sold

```

### Compute Sold Per Day Column & Projected Revenue Figures
```{r}
poor_mans_filtered$sold_per_day = poor_mans_filtered$number_sold/poor_mans_filtered$days_on_menu
poor_mans_filtered$projected_revenue = poor_mans_filtered$Price.Sold*poor_mans_filtered$sold_per_day*365
poor_mans_filtered

medallions_filtered$sold_per_day = medallions_filtered$number_sold/medallions_filtered$days_on_menu
medallions_filtered$projected_revenue = medallions_filtered$Price.Sold*medallions_filtered$sold_per_day*365
medallions_filtered

famous_pepper_filtered$sold_per_day = famous_pepper_filtered$number_sold/famous_pepper_filtered$days_on_menu
famous_pepper_filtered$projected_revenue = famous_pepper_filtered$Price.Sold*famous_pepper_filtered$sold_per_day*365
famous_pepper_filtered
```

### Calculate Elasticity for POS Collected Data
```{r}
#Low Tier: Poor Man's Pepper Steak 
low_q_delta = (max(poor_mans_filtered$sold_per_day) - min(poor_mans_filtered$sold_per_day))/((max(poor_mans_filtered$sold_per_day)+min(poor_mans_filtered$sold_per_day))/2)

low_p_delta = (max(poor_mans_filtered$Price.Sold) - min(poor_mans_filtered$Price.Sold))/((max(poor_mans_filtered$Price.Sold)+min(poor_mans_filtered$Price.Sold))/2)

low_tier_elastciity = low_q_delta/low_p_delta
low_tier_elastciity

#Medium Tier: Medallions of Filet Mignon
med_q_delta = (max(medallions_filtered$sold_per_day) - min(medallions_filtered$sold_per_day))/((max(medallions_filtered$sold_per_day)+min(medallions_filtered$sold_per_day))/2)

med_p_delta = (max(medallions_filtered$Price.Sold) - min(medallions_filtered$Price.Sold))/((max(medallions_filtered$Price.Sold)+min(medallions_filtered$Price.Sold))/2)

medium_tier_elastcity = med_q_delta/med_p_delta
medium_tier_elastcity

#High Tier: Dal Rae Famous Pepper Steak 12oz
high_q_delta = (max(famous_pepper_filtered$sold_per_day) - min(famous_pepper_filtered$sold_per_day))/((max(famous_pepper_filtered$sold_per_day)+min(famous_pepper_filtered$sold_per_day))/2)

high_p_delta = (max(famous_pepper_filtered$Price.Sold) - min(famous_pepper_filtered$Price.Sold))/((max(famous_pepper_filtered$Price.Sold)+min(famous_pepper_filtered$Price.Sold))/2)

high_tier_elasticity = high_q_delta/high_p_delta
high_tier_elasticity

```

### Graph POS Data Product Demand Curves 
```{r}
low_lm_model = lm(sold_per_day ~ Price.Sold, data=poor_mans_filtered)
summary(low_lm_model)
#plot(low_lm_model)

med_lm_model = lm(sold_per_day ~ Price.Sold, data=medallions_filtered)
summary(med_lm_model)
#plot(med_lm_model)

high_lm_model = lm(sold_per_day ~ Price.Sold, data=famous_pepper_filtered)
summary(high_lm_model)
#plot(high_lm_model)

library("ggplot2")

ggplot(data = poor_mans_filtered, aes(x = sold_per_day, y = Price.Sold)) +
geom_point() + geom_smooth(method = "lm", se=FALSE) +
xlab("Sold per Day") +
ylab("Price Points") +
ggtitle("POS Data: Poor Man's Pepper Steak")

ggplot(data = medallions_filtered, aes(x = sold_per_day, y = Price.Sold)) +
geom_point() + geom_smooth(method = "lm", se=FALSE) +
xlab("Sold per Day") +
ylab("Price Points") +
ggtitle("POS Data: Medallions of Filet Mignon")

ggplot(data = famous_pepper_filtered, aes(x = sold_per_day, y = Price.Sold)) +
geom_point() + geom_smooth(method = "lm", se=FALSE) +
xlab("Sold per Day") +
ylab("Price Points") +
ggtitle("POS Data: Dal Rae Famous Pepper Steak 12oz")
```
### Clean Data For Survey Demand Curves
```{r}
survey_data <- read_excel("~/Desktop/Dal Rae CSVs/Dal Rae Restaurant Customer Survey (Responses).xlsx")

poor_mans_repsonses = subset(survey_data, select = c('Poor Mans 30 Discount','Poor Mans 20 Discount','Poor Mans 10 Discount','Poor Mans 5 Discount','Poor Mans 5 Increase', 'Poor Mans 10 Increase','Poor Mans 20 Increase','Poor Mans 30 Increase'))

medallions_responses = subset(survey_data, select = c('Medallions 30 Discount','Medallions 20 Discount','Medallions 10 Discount','Medallions 5 Discount','Medallions 5 Increase', 'Medallions 10 Increase','Medallions 20 Increase','Medallions 30 Increase'))

famous_pepper_responses = subset(survey_data, select = c('Famous Pepper 30 Discount','Famous Pepper 20 Discount','Famous Pepper 10 Discount','Famous Pepper 5 Discount','Famous Pepper 5 Increase', 'Famous Pepper 10 Increase','Famous Pepper 20 Increase','Famous Pepper 30 Increase'))


frequency_converter <- function(column) {
  for(i in 1:length(column)) {
    if (column[i] == 'would not order') {
      column[i] = 0
      } else if (column[i] == 'every year') {
      column[i] = 1/365
      } else if (column[i] == 'every two months') {
      column[i] = 6/365
      } else if (column[i] == 'every month') {
      column[i] = 12/365
      } else if (column[i] == 'every two weeks') {
      column[i] = 26/365
      } else if (column[i] == 'every week') {
      column[i] = 52/365
      } else if (column[i] == '2x per week') {
      column[i] = 104/365
      } else {
      column[i] = 156/365
    }
  } 
return(column)
}

#Changes Frequency Values
poor_mans_repsonses$`Poor Mans 30 Discount` = frequency_converter(poor_mans_repsonses$`Poor Mans 30 Discount`)
poor_mans_repsonses$`Poor Mans 20 Discount` = frequency_converter(poor_mans_repsonses$`Poor Mans 20 Discount`)
poor_mans_repsonses$`Poor Mans 10 Discount` = frequency_converter(poor_mans_repsonses$`Poor Mans 10 Discount`)
poor_mans_repsonses$`Poor Mans 5 Discount` = frequency_converter(poor_mans_repsonses$`Poor Mans 5 Discount`)
poor_mans_repsonses$`Poor Mans 5 Increase` = frequency_converter(poor_mans_repsonses$`Poor Mans 5 Increase`)
poor_mans_repsonses$`Poor Mans 10 Increase` = frequency_converter(poor_mans_repsonses$`Poor Mans 10 Increase`)
poor_mans_repsonses$`Poor Mans 20 Increase` = frequency_converter(poor_mans_repsonses$`Poor Mans 20 Increase`)
poor_mans_repsonses$`Poor Mans 30 Increase` = frequency_converter(poor_mans_repsonses$`Poor Mans 30 Increase`)

medallions_responses$`Medallions 30 Discount` = frequency_converter(medallions_responses$`Medallions 30 Discount`)
medallions_responses$`Medallions 20 Discount` = frequency_converter(medallions_responses$`Medallions 20 Discount`)
medallions_responses$`Medallions 10 Discount` = frequency_converter(medallions_responses$`Medallions 10 Discount`)
medallions_responses$`Medallions 5 Discount` = frequency_converter(medallions_responses$`Medallions 5 Discount`)
medallions_responses$`Medallions 5 Increase` = frequency_converter(medallions_responses$`Medallions 5 Increase`)
medallions_responses$`Medallions 10 Increase` = frequency_converter(medallions_responses$`Medallions 10 Increase`)
medallions_responses$`Medallions 20 Increase` = frequency_converter(medallions_responses$`Medallions 20 Increase`)
medallions_responses$`Medallions 30 Increase` = frequency_converter(medallions_responses$`Medallions 30 Increase`)

famous_pepper_responses$`Famous Pepper 30 Discount` = frequency_converter(famous_pepper_responses$`Famous Pepper 30 Discount`)
famous_pepper_responses$`Famous Pepper 20 Discount` = frequency_converter(famous_pepper_responses$`Famous Pepper 20 Discount`)
famous_pepper_responses$`Famous Pepper 10 Discount` = frequency_converter(famous_pepper_responses$`Famous Pepper 10 Discount`)
famous_pepper_responses$`Famous Pepper 5 Discount` = frequency_converter(famous_pepper_responses$`Famous Pepper 5 Discount`)
famous_pepper_responses$`Famous Pepper 5 Increase` = frequency_converter(famous_pepper_responses$`Famous Pepper 5 Increase`)
famous_pepper_responses$`Famous Pepper 10 Increase` = frequency_converter(famous_pepper_responses$`Famous Pepper 10 Increase`)
famous_pepper_responses$`Famous Pepper 20 Increase` = frequency_converter(famous_pepper_responses$`Famous Pepper 20 Increase`)
famous_pepper_responses$`Famous Pepper 30 Increase` = frequency_converter(famous_pepper_responses$`Famous Pepper 30 Increase`)

#Changes DataType of Each Column to Numeric

# Function to change data type of a column
poor_mans_repsonses = as.data.frame(lapply(poor_mans_repsonses, as.numeric))
medallions_responses = as.data.frame(lapply(medallions_responses, as.numeric))
famous_pepper_responses = as.data.frame(lapply(famous_pepper_responses, as.numeric))

#Computes Averages of Each Column
poor_mans_average_freq = colMeans(poor_mans_repsonses)
medallions_average_freq = colMeans(medallions_responses)
famous_pepper_average_freq = colMeans(famous_pepper_responses)
```
### Graph Survey Demand Curves
```{r}
library(dplyr)

price_levels = c(0.7, 0.8, 0.9, 0.95, 1.05, 1.1, 1.2, 1.3)

price_poor = 22.75
price_poor_vec = round(price_levels * price_poor, 2)

price_medallions = 44
price_medallions_vec = round(price_levels * price_medallions, 2)

price_famous_pepper = 69.50
price_famous_pepper_vec = round(price_levels * price_famous_pepper, 2)
  
poor_df = data.frame(price_poor_vec, poor_mans_average_freq)
poor_df$days_per_year = poor_df$poor_mans_average_freq * 365
poor_df <- poor_df %>% rename("prices" = "price_poor_vec")

medallions_df = data.frame(price_medallions_vec, medallions_average_freq)
medallions_df$days_per_year = medallions_df$medallions_average_freq * 365
medallions_df <- medallions_df %>% rename("prices" = "price_medallions_vec")

famous_pepper_df = data.frame(price_famous_pepper_vec, famous_pepper_average_freq)
famous_pepper_df$days_per_year = famous_pepper_df$famous_pepper_average_freq * 365
famous_pepper_df <- famous_pepper_df %>% rename("prices" = "price_famous_pepper_vec")

poor_df
medallions_df
famous_pepper_df

#Creates Linear Regression Models
med_survey_model = lm(days_per_year ~ prices, data=medallions_df)
#summary(med_lm_model)
#plot(med_lm_model)

high_survey_model = lm(days_per_year ~ prices, data=famous_pepper_df)
#summary(high_lm_model)
#plot(high_lm_model)

#Graphs The Frequency of Each Entree
ggplot(data = poor_df, aes(x = days_per_year, y = prices)) +
geom_point() + geom_smooth(method = "lm", se=FALSE) +
xlab("Average Frequency (Days Per Year)") +
ylab("Price Points") +
ggtitle("Survey Data: Poor Man's Pepper Steak") + scale_y_continuous(breaks= price_poor_vec)

ggplot(data = medallions_df, aes(x = days_per_year, y = prices)) +
geom_point() + geom_smooth(method = "lm", se=FALSE) +
xlab("Average Frequency (Days Per Year)") +
ylab("Price Points") +
ggtitle("Survey Data: Medallions of Filet Mignon") + scale_y_continuous(breaks= price_medalllions_vec)

ggplot(data = famous_pepper_df, aes(x = days_per_year, y = prices)) +
geom_point() + geom_smooth(method = "lm", se=FALSE) +
xlab("Average Frequency (Days Per Year)") +
ylab("Price Points") +
ggtitle("Survey Data: Dal Rae Famous Pepper Steak") + scale_y_continuous(breaks= price_famous_pepper_vec)

```

### Calculate Elasticity for Survey Data
```{r}
#Poor Man's Pepper Steak
low_q_delta = (max(poor_df$days_per_year) - min(poor_df$days_per_year))/((max(poor_df$days_per_year)+min(poor_df$days_per_year))/2)

low_p_delta = (max(poor_df$prices) - min(poor_df$prices))/((max(poor_df$prices)+min(poor_df$prices))/2)

low_tier_elastciity = low_q_delta/low_p_delta
low_tier_elastciity

#Medallions of Filet Mignon
med_q_delta = (max(medallions_df$days_per_year) - min(medallions_df$days_per_year))/((max(medallions_df$days_per_year)+min(medallions_df$days_per_year))/2)

med_p_delta = (max(medallions_df$prices) - min(medallions_df$prices))/((max(medallions_df$prices)+min(medallions_df$prices))/2)

med_tier_elastciity = med_q_delta/med_p_delta
med_tier_elastciity

#Dal Rae Famous Pepper Steak
high_q_delta = (max(famous_pepper_df$days_per_year) - min(famous_pepper_df$days_per_year))/((max(famous_pepper_df$days_per_year)+min(famous_pepper_df$days_per_year))/2)

high_p_delta = (max(famous_pepper_df$prices) - min(famous_pepper_df$prices))/((max(famous_pepper_df$prices)+min(famous_pepper_df$prices))/2)
high_tier_elastciity = high_q_delta/high_p_delta
high_tier_elastciity
```

### Triangulation of Survey Data + POS Data for Price Increase Quantity Estimates
```{r}
#Price Definition
new_data_med = data.frame(Price.Sold = 55)
new_data_pepper = data.frame(Price.Sold = 82)

new_data_med_survey = data.frame(prices = 55)
new_data_pepper_survey = data.frame(prices = 82)

#POS Prediction Medallions
#percent_delta = -(1 - (predict(med_lm_model, new_data_med)/medallions_filtered$sold_per_day[8]))
#percent_delta

#Survey Data Prediction Medallions
percent_delta = -(1- (predict(med_survey_model, new_data_med_survey)/((medallions_df$days_per_year[4] + medallions_df$days_per_year[5])/2)))
predict(med_survey_model, new_data_med_survey)
((medallions_df$days_per_year[4] + medallions_df$days_per_year[5])/2)
percent_delta
                  

#Triangulate Medallions


#POS Prediction Famous Pepper
#predict(high_lm_model, new_data_pepper)

#Survey Prediction Famous Pepper
#predict(high_survey_model, new_data_pepper_survey)
```


